R for Clinical Study Reports and Submission

### 2023-01-16 preface.Rmd

We recommend using the
[R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf)
to organize all the A&R-related source code and documentation for a clinical trial A&R project.
The R package folder structure is well defined and widely used in the R community through repositories (e.g., CRAN).

@@

>>>>

We recommend using the
[R package folder structure](https://github.com/rstudio/cheatsheets/blob/main/package-development.pdf)
to organize all the A&R-related source code and documentation for a clinical trial A&R project.
The R package folder structure is well defined and widely used in the R community through repositories (e.g., CRAN).

@@

```{block2, type="flushright", html.tag="p"}
Yilong Zhang, Nan Xiao, Keaven Anderson, Yalin Zhu

December 2022
```

### 2023-01-16 slides/r4csr-rstudio.Rmd

- R is widely used in visualization
  - [SafetyGraphics](https://github.com/SafetyGraphics/safetyGraphics)
  - [forestly](https://elong0527.github.io/forestly/articles/example.html#construct-interactive-forest-plot-1): [Interactive forest plot for DMC safety monitoring in clinical trials](https://rinpharma.com/publication/rinpharma_194/)

>>>>

- R is widely used in visualization
  - [SafetyGraphics](https://github.com/SafetyGraphics/safetyGraphics)
  - [forestly](https://elong0527.github.io/forestly/articles/example.html#construct-interactive-forest-plot-1): [Interactive forest plot for DMC safety monitoring in clinical trials](https://rinpharma.com/publication/rinpharma_206/)

### 2023-01-16 tlf-overview.Rmd

In clinical trials, a critical step is to submit trial results to regulatory agencies.
[Electronic Common Technical Document (eCTD)](https://en.wikipedia.org/wiki/Electronic_common_technical_document)
has become a worldwide regulatory submission standard format.
For example, the United States Food and Drug Administration (US FDA) requires
new drug applications and biologics license applications
[must be submitted using the eCTD format](https://www.fda.gov/drugs/electronic-regulatory-submission-and-review/electronic-common-technical-document-ectd).
The Clinical Data Interchange Standards Consortium (CDISC) provides a
[pilot project following ICH E3 guidance](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse).

@@

A typical CSR contains full details on the methods and results of an individual clinical study.
In support of the statistical analysis, a large number of tables, listings, and figures are
incorporated into the main text and appendices.
In the CDISC pilot project, an
[example CSR](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf)
is also provided. If you are interested in more examples of clinical study reports,
you can go to the [European Medicines Agency (EMA) clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

@@

We used publicly available CDISC pilot
[study data located in the CDISC Bitbucket repository](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/datasets/cdiscpilot01/analysis/adam/datasets).

@@

tidyverse is a collection of R packages to simplify the workflow to manipulate,
visualize and analyze data in R.
Those R packages share
[the tidy tools manifesto](https://cran.r-project.org/web/packages/tidyverse/vignettes/manifesto.html)
and are easy to use for interactive data analysis.

RStudio provided outstanding [cheatsheets](https://www.rstudio.com/resources/cheatsheets/)
and [tutorials](https://github.com/rstudio-education/remaster-the-tidyverse) for tidyverse.

>>>>

In clinical trials, a critical step is to submit trial results to regulatory agencies.
[Electronic Common Technical Document (eCTD)](https://en.wikipedia.org/wiki/Electronic_common_technical_document)
has become a worldwide regulatory submission standard format.
For example, the United States Food and Drug Administration (US FDA) requires
new drug applications and biologics license applications
[must be submitted using the eCTD format](https://www.fda.gov/drugs/electronic-regulatory-submission-and-review/electronic-common-technical-document-ectd).
The Clinical Data Interchange Standards Consortium (CDISC) provides a
[pilot project following ICH E3 guidance](https://github.com/cdisc-org/sdtm-adam-pilot-project).

@@

A typical CSR contains full details on the methods and results of an individual clinical study.
In support of the statistical analysis, a large number of tables, listings, and figures are
incorporated into the main text and appendices.
In the CDISC pilot project, an
[example CSR](https://github.com/cdisc-org/sdtm-adam-pilot-project/blob/master/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf)
is also provided. If you are interested in more examples of clinical study reports,
you can go to the [European Medicines Agency (EMA) clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

@@

We used publicly available CDISC pilot
[study data located in the CDISC GitHub repository](https://github.com/cdisc-org/sdtm-adam-pilot-project/tree/master/updated-pilot-submission-package/900172/m5/datasets/cdiscpilot01/analysis/adam/datasets).

@@

tidyverse is a collection of R packages to simplify the workflow to manipulate,
visualize and analyze data in R.
Those R packages share
[the tidy tools manifesto](https://tidyverse.tidyverse.org/articles/manifesto.html)
and are easy to use for interactive data analysis.

RStudio provided outstanding [cheatsheets](https://posit.co/resources/cheatsheets/)
and [tutorials](https://github.com/rstudio-education/remaster-the-tidyverse) for tidyverse.
